---
title: "ST558 - Homework 4: Hodge podging with purrr"
authors: Laraib Azmat
date: June 20, 2024
format: html
---

```{r}
library(httr)
library(jsonlite)
library(tidyverse)
```


### Task 1: Conceptual Questions

1.    What is the purpose of the `lapply()` function? What is the equivalent `purrr` function?
 - `lapply()` applies a given function to a list  

2.    Suppose we have a list called my_list. Each element of the list is a numeric data frame (all columns are numeric). We want use `lapply()` to run the code cor(numeric_matrix, method = "kendall") on each element of the list. Write code to do this below! (I’m really trying to ask you how you specify `method = "kendall"` when calling `lapply()`). 

  -   `lapply(X = my_list, FUN = function(x) {cor(x, method = "kendall")})`
  
3.    What are two advantages of using `purrr` functions instead of the BaseR apply family?

  -   Type stable functions and first argument is always the data 
  
4.    What is a side-effect function?

  -   Does something other than return a value 
  
5.    Why can you name a variable `sd` in a function and not cause any issues with the `sd` function?

  -   The function exists in a different namespace than the variable 
  

### Task 2: Writing R Functions
```{r}
#setting up trial data to test function
set.seed(5)
n <- 50
x <- runif(n)
response <- 2 + 10*x + rnorm(n)
predict <- predict(lm(response ~ x), data.frame(x))
```

```{r}
#running basic computation for RMSE that I will later put into the function, also getting an idea of the value
sqrt(mean((response - predict)^2))
```

```{r}
#creating a function to run the RMSE computation
getRMSE <- function(x, y, ...){
  stat <- sqrt(mean((x - y)^2, ...))
  return(stat = stat)
}
```

```{r}
#testing the function
getRMSE(response, predict)
```

```{r}
#creating objects from task 1, #2
set.seed(10)
n <- 100
x <- runif(n)
resp2 <- 3 + 10*x + rnorm(n)
pred2 <- predict(lm(resp2 ~ x), data.frame(x))
```

```{r}
#finding the correct results first 
sqrt(mean((resp2 - pred2)^2))
```

```{r}
#running the RMSE function and comparing it to previous results
getRMSE(resp2, pred2)
```
```{r}
#adding NA values to the data 
resp2[55:56] <- NA_real_

#running RMSE function without specifying NA values
getRMSE(resp2, pred2)

#testing the additional arguments by specifying the NA values 
getRMSE(resp2, pred2, na.rm = TRUE)
```


```{r}
#repeating computation step with MAE 
abs(mean(response - predict))
```

```{r}
#creating a function for MAE
getMAE <- function(x, y, ...){
  stat <- abs(mean(x - y, ...))
  return(stat)
}
```

```{r}
#testing the function
getMAE(response, predict)
```

```{r}
#Task 2, #4 data
set.seed(10)
n <- 100
x <- runif(n)
resp4 <- 3 + 10*x + rnorm(n)
pred4 <- predict(lm(resp4 ~ x), data.frame(x))
```

```{r}
#finding the correct results first 
abs(mean(resp4 - pred4))
```

```{r}
#running the MAE function 
getMAE(resp4, pred4)
```
```{r}
#adding NA values to the data 
resp4[42:43] <- NA_real_

#running the MAE function without specifying NA values
getMAE(resp4, pred4)

#testing the additional arguments by specifying the NA values 
getMAE(resp4, pred4, na.rm = TRUE)
```

```{r}
stat_wrapper <- function(x, y, ...){
  if(!(is.vector(x) & is.atomic(x) & is.numeric(x))){
    stop("Response variable is invalid")
  }
  if(!(is.vector(y) & is.atomic(y) & is.numeric(y))){
    stop("Predictor variable is invalid")
  }
  Stats <- list()
  Stats$RMSE <- getRMSE(x, y, ...)
  Stats$MAE <- getMAE(x, y, ...)
  return(Stats)
}
```

```{r}
#Task 2, #6 data
set.seed(10)
n <- 100
x <- runif(n)
resp6 <- 3 + 10*x + rnorm(n)
pred6 <- predict(lm(resp6 ~ x), data.frame(x))
```

```{r}
#running base functions to get the correct values
getRMSE(resp6, pred6)
getMAE(resp6, pred6)
```

```{r}
#running the wrapper function 
stat_wrapper(resp6, pred6)
```

```{r}
#adding NA values to the data
resp6[22:23] <- NA_real_

#running the wrapper function without specifying NA values
stat_wrapper(resp6, pred6)

#testing the additional arguments by specifying the NA values 
stat_wrapper(resp6, pred6, na.rm = TRUE)
```

```{r error=TRUE}
#testing with incorrect data type 
stat_wrapper(iris)
```

### Task 3: Querying an API and a Tidy-Style Function

```{r}
#adding hidden api key file
api_key <- source("api_key.R")
```

```{r}
#grabbing some articles using personal API key
bio_api <- httr::GET("https://newsapi.org/v2/everything?q=biofuel", 
                     add_headers(Authorization = (api_key$value)))
```

```{r}
#parsing the API info
bio_parsed <- bio_api$content |>
  rawToChar() |>
  fromJSON()
str(bio_parsed, max.level = 1)
```

```{r}
#turning the dataframe into a tibble 
bio_info <- bio_parsed$articles |>
  as_tibble()
bio_info
```

```{r}
news_articles <- function(subject, time_frame_start, time_frame_end, key = api_key$value){
  URL <- paste("https://newsapi.org/v2/everything?",
               subject,
               "&from=",
               time_frame_start,
               "&to=",
               time_frame_end,
               "&language=en&apiKey=", 
               key, 
               sep = "")
  temp_api <- GET(URL)
  
  temp_parse <- temp_api$content |>
    rawToChar() |>
    fromJSON() |>
    as_tibble()
  return(temp_parse)
}
```

```{r}
news_articles("biofuel", "2024-06-29", "2024-01-01")
news_articles("apple", "2024-06-29", "2024-01-01")
```

  -   I am not sure why the request isn't working in this case. 